\section{介词}

\subsection{介词概述}

德语中的介词对其后词语的格有明确的要求，因此需要记忆介词和其支配的格。同时德语的介词与英语的也不能总是一一对应，也有很多固定搭配需要记忆。

德语的介词大多放在名词之前，但也有少部分可以放在名词之后(部分是固定用法)。也有部分介词是由两部分组成，将名词放在之间使用。

德语的介词连用时，名词的格按照最近的介词确定。有时介词也可以发生合并。



\subsection{介词与冠词的合写}

\begin{table}[htbp]
    \caption{常见的介词-冠词合写}
    \label{tab:pa-contraction}
    \centering
\begin{tblr}{
    width=\textwidth,
    colspec={rcl||rcl},
    column{1,4}={rightsep=3pt},
    column{2,5}={colsep=0pt,mode=math},
    column{3,6}={leftsep=3pt},
}
an das    & \Rightarrow & ans    & zu der     & \Rightarrow & zur     \\
auf das   & \Rightarrow & aufs   & an dem     & \Rightarrow & am      \\
durch das & \Rightarrow & durchs & bei dem    & \Rightarrow & beim    \\
für das   & \Rightarrow & fürs   & hinter dem & \Rightarrow & hinterm \\
in das    & \Rightarrow & ins    & in dem     & \Rightarrow & im      \\
über das  & \Rightarrow & übers  & unter dem  & \Rightarrow & unterm  \\
um das    & \Rightarrow & ums    & von dem    & \Rightarrow & vom     \\
unter das & \Rightarrow & unters & vor dem    & \Rightarrow & vorm    \\
vor das   & \Rightarrow & vors   & zu dem     & \Rightarrow & zum     \\
\end{tblr}
\end{table}

可以发现，基本所有的合写都是在介词后添加冠词的最后一个字母，除了 an dem, in dem, von dem 这三个合写，把介词的最后一个字母直接替换为本应附加的字母。应该也是因为发音和书写方便的原因。

\subsection{介词表}

列出了所有介词和英文中与其对应的单词，标*指该介词可以后置使用。

A、D、G 代表该用法下介词支配的格

\clearpage

\begin{longtblr}[
    caption = {介词表},
    entry = {Short Caption},
    label = {tab:prepositions},
]{
    width=\linewidth,
    rowhead=1,
    colspec={lllX},
    columns={l,colsep=3pt},
    column{1}={wd=2cm},
    column{3}={wd=3cm},
}
    \SetCell[c=2]{c} \textbf{Prep. and Case} & & \textbf{Equivalents} & \textbf{Sample usage and notes} \\
    \hline
    \SetCell[r=2]{m} betreffend* & \SetCell[r=2]{m} A & \SetCell[r=2]{m} with reference to, concerning & Ihre Anfrage \textbf{betreffend}… \\
    & & & regarding your inquiry… \\
    \hline
    \SetCell[r=4]{m} bis & \SetCell[r=2]{m} A & \SetCell[r=2]{m} until & Zähle von 1 \textbf{bis} 10. \\
    &       &       & Count from 1 to 10. \\
    \hline
    & \SetCell[r=2]{m} A & \SetCell[r=2]{m} by & \textbf{Bis} wann mussen wir fertig sein? \\
    &       &       & By when do we have to be finished? \\
    \hline
    \SetCell[r=8]{m} durch & \SetCell[r=2]{m} A & \SetCell[r=2]{m} through & Es bringt mich \textbf{durch} den Tag. \\
    &       &       & It gets me through the day. \\
    \hline
    & \SetCell[r=2]{m} A & \SetCell[r=2]{m} by & \textbf{durch} einen Anwalt vertreten werden \\
    &       &       & to be represented by a lawyer \\
    \hline
    & \SetCell[r=2]{m} A & \SetCell[r=2]{m} because of & \textbf{Durch} den Regen war das Picknick ein Fiasko. \\
    &       &       & Because of the rain, the picnic was a fiasco. \\
    \hline
    & \SetCell[r=2]{m} A & \SetCell[r=2]{m} via & Ich habe es \textbf{durch} Freunde erfahren. \\
    &       &       & I found out through/from friends. \\
    \hline
    \SetCell[r=2]{m} für & \SetCell[r=2]{m} A & \SetCell[r=2]{m} for & eine Nachricht \textbf{für} dich \\
    &       &       & a message for you \\
    \hline
    \SetCell[r=6]{m} entlang* & \SetCell[r=6]{m} A & \SetCell[r=6]{m} along (side) & den Strand \textbf{entlang} laufen \\
    &       &       & to walk along the beach \\
    &       &       & \textcolor{codegray}{[accusative when postpositional]}\\
    \hline
    &       &       & die Häuser \textbf{entlang} der Straße \\
    &       &       & the houses along the street \\
    &       &       & \textcolor{codegray}{[but usually genitive when prepositional]}\\
    \hline
    \SetCell[r=6]{m} gegen & \SetCell[r=4]{m} A & \SetCell[r=4]{m} against & „\textbf{Gegen} die Wand” \\
    &       &       & Against the Wall \textcolor{codegray}{[movie title]} \\
    \hline
    &       &       & Alle sind \textbf{gegen} mich. \\
    &       &       & Everyone's against me. \\
    \hline
    & \SetCell[r=2]{m} A & \SetCell[r=2]{m} around/about [a time] & \textbf{Gegen} Mitternacht gingen wir ins Bett. \\
    &       &       & We went to bed around midnight. \\
    \hline
    \SetCell[r=4]{m} ohne & \SetCell[r=4]{m} A & \SetCell[r=4]{m} without & \textbf{ohne} Zweifel \\
    &       &       & without [a] doubt \\
    \hline
    &       &       & \textbf{Ohne} dich kann ich nicht leben. \\
    &       &       & I can't live without you. \\
    \hline
    \SetCell[r=4]{m} um & \SetCell[r=2]{m} A & \SetCell[r=2]{m} around & \textbf{um} den Block laufen \\*
    &       &       & to walk around the block \\*
    \hline
    & \SetCell[r=2]{m} A & \SetCell[r=2]{m} at [a time] & Wir treffen uns \textbf{um} 14:00 Uhr. \\*
    &       &       & We're meeting at 2 pm. \\*
    \hline
    \SetCell[r=3]{m} wider & \SetCell[r=3]{m} A & \SetCell[r=3]{m} against & \textbf{wider} meinen Willen \\
    &       &       & against my will \\
    &       &       & \textcolor{codegray}{[more idiomatic / literary than "gegen"]}\\
    \hline
    \SetCell[r=2]{m} ab & \SetCell[r=2]{m} D & \SetCell[r=2]{m} from [a point in time] & \textbf{ab} dem ersten Schultag \\
    &       &       & from the first day of school \\
    \hline
    \SetCell[r=6]{m} aus & \SetCell[r=2]{m} D & \SetCell[r=2]{m} from [origin or source] & Ich komme \textbf{aus} den USA. \\
    &       &       & I come from the US. \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} from (out of) & Tomaten \textbf{aus} der Dose \\
    &       &       & tomatoes out of the can \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} from (made of) & \textbf{aus} Metall \\
    &       &       & made from/of metal \\
    \hline
    \SetCell[r=4]{m} außer & \SetCell[r=2]{m} D & \SetCell[r=2]{m} except for, besides, apart from & keiner \textbf{außer} mir \\
    &       &       & no one but me \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} out of & \textbf{außer} Kontrolle \\
    &       &       & out of control \\
    \hline
    \SetCell[r=4]{m} bei & \SetCell[r=2]{m} D & \SetCell[r=2]{m} at & Wasser kocht \textbf{bei} hundert Grad. \\
    &       &       & Water boils at 100 degrees. \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} with / (at the house of) & Ich bleibe \textbf{bei} dir \\
    &       &       & I'll stay with you / at your place \\
    % \hline
    \pagebreak
    \SetCell[r=6]{m} entgegen* & \SetCell[r=2]{m} D & \SetCell[r=2]{m} towards & Sie lief mir \textbf{entgegen}. \\
    &       &       & She ran towards me. \\
    \hline
    & \SetCell[r=4]{m} D & \SetCell[r=4]{m} contrary to, notwithstanding & Dem steht nichts \textbf{entgegen}. \\
    &       &       & There's nothing to stop it from happening. \\
    \hline
    &       &       & \textbf{entgegen} den Erwartungen... \\
    &       &       & contrary to expectations... \\
    \hline
    \SetCell[r=2]{m} entsprechend* & \SetCell[r=2]{m} D & \SetCell[r=2]{m} according to, corresponding to & den Umständen \textbf{entsprechend}... \\
    &       &       & considering the circumstances... \\
    \hline
    \SetCell[r=4]{m} gegenüber* & \SetCell[r=2]{m} D & \SetCell[r=2]{m} across from & \textbf{gegenüber} der Strasse \\*
    &       &       & across the street \\*
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} to (a person) & Mir \textbf{gegenüber} hat er nichts darüber gesagt. \\*
    &       &       & He didn’t say anything to me about it. \\*
    \hline
    \SetCell[r=4]{m} gemäß* & \SetCell[r=4]{m} D & \SetCell[r=4]{m} according to, in accordance with & \textbf{gemäß} Paragraf 27 des Strafgesetzbuches ... \\
    &       &       & {according to Paragraph 27 of \\ the criminal law code ...} \\
    \hline
    &       &       & unseren Instinkten \textbf{gemäß}… \\
    &       &       & according to our instincts… \\
    \hline
    \SetCell[r=3]{m} laut & \SetCell[r=3]{m} D & \SetCell[r=3]{m} according to & \textbf{laut} dem Pressesprecher... \\
    &       &       & according to the press spokesman... \\
    &       &       & \textcolor{codegray}{[can also be genitive (laut des Presseprechers)]}\\
    \hline
    \SetCell[r=7]{m} mit & \SetCell[r=3]{m} D & \SetCell[r=3]{m} with & Bier \textbf{mit} Cola \\
    &       &       & beer with cola \\
    &       &       & \textcolor{codegray}{[popular German drink]}\\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} over [time] & \textbf{mit} den Jahren \\
    &       &       & over the years \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} at [an age] & ein Kind \textbf{mit} 2 Jahren \\
    &       &       & a two year old child \\
    % \hline
    \pagebreak
    \SetCell[r=4]{m} (mit)samt & \SetCell[r=4]{m} D & \SetCell[r=4]{m} including, together with & \textbf{samt} Familie \\
    &       &       & with one's whole family \\
    \hline
    &       &       & \textbf{samt} Gepäck \\
    &       &       & with all one's baggage \\
    \hline
    \SetCell[r=7]{m} nach* & \SetCell[r=2]{m} D & \SetCell[r=2]{m} after & \textbf{nach} der Arbeit/Schule \\
    &       &       & after work/school \\
    \hline
    & \SetCell[r=3]{m} D & \SetCell[r=3]{m} according to & \textbf{nach} meiner Meinung  \\
    &       &       & according to (in) my opinion \\
    &       &       & \textcolor{codegray}{[or meiner Meinung \textbf{nach}]}\\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} to & Ich fliege \textbf{nach} Deutschland \\
    &       &       & I'm flying to Germany \\
    \hline
    \SetCell[r=2]{m} nahe & \SetCell[r=2]{m} D & \SetCell[r=2]{m} near & \textbf{nahe} der Natur \\
    &       &       & close to nature \\
    \hline
    \SetCell[r=4]{m} seit & \SetCell[r=2]{m} D & \SetCell[r=2]{m} since [point in time] & \textbf{seit} September \\
    &       &       & since September \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} for [time period] & \textbf{seit} langer Zeit \\
    &       &       & for a long time \\
    \hline
    \SetCell[r=6]{m} von & \SetCell[r=2]{m} D & \SetCell[r=2]{m} from [immediate origin] & Ich komme \textbf{von} der Arbeit. \\*
    &       &       & I'm coming from work. \\*
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} of & die Königin \textbf{von} England \\*
    &       &       & the Queen of England \\*
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} by & \textbf{von} Hand gemacht \\*
    &       &       & made by hand \\*
    \hline
    \SetCell[r=4]{m} zu & \SetCell[r=2]{m} D & \SetCell[r=2]{m} to & Ich gehe \textbf{zu}m Arzt. \\
    &       &       & I'm going to the doctor. \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} at & \textbf{zu} diesem Zeitpunkt \\
    &       &       & at this point in time \\
    % \hline
    \pagebreak
    \SetCell[r=6]{m} an & \SetCell[r=4]{m} A & \SetCell[r=4]{m} at, on & Setz dich \textbf{an} den Tisch. \\
    &       &       & Sit down at the table. \\
    \hline
    &       &       & ein Bild \textbf{an} die Wand hängen \\
    &       &       & to hang a picture on the wall \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} at, on & Das Bild hängt \textbf{an} der W\textbf{an}d. \\
    &       &       & The picture hangs (is hanging) on the wall. \\
    \hline
    \SetCell[r=7]{m} auf & \SetCell[r=4]{m} A & \SetCell[r=4]{m} on, up & \textbf{auf} den Berg steigen \\
    &       &       & to climb up the mountain \\
    \hline
    &       &       & eine Vase \textbf{auf} den Tisch stellen \\
    &       &       & to set a vase down on the table \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} on, up & Die Vase steht \textbf{auf} dem Tisch. \\
    &       &       & The vase is (standing) on the table. \\
    \hline
    & \SetCell[c=3]{l,m,\textwidth-3cm} {The word \textbf{auf} implies a position on top of something; \textbf{an} is more like "beside" or suspended from.}\\
    \hline
    \SetCell[r=4]{m} hinter & \SetCell[r=2]{m} A & \SetCell[r=2]{m} behind & ein Blick \textbf{hinter} die Kulissen \\
    &       &       & a glimpse behind the scenes \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} behind & Was passiert \textbf{hinter} den Kulissen? \\
    &       &       & What's happening behind the scenes? \\
    \hline
    \SetCell[r=4]{m} in & \SetCell[r=2]{m} A & \SetCell[r=2]{m} into, to & Die Kinder gehen \textbf{in} die Schule. \\
    &       &       & The kids are going to school. \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} in & Was hast du \textbf{in} der Schule gelernt? \\
    &       &       & What did you learn in school? \\
    \hline
    \SetCell[r=4]{m} neben & \SetCell[r=2]{m} A & \SetCell[r=2]{m} next to (adjacent) & Komm, setz dich \textbf{neben} mich. \\
    &       &       & Come sit down next to me. \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} next to (adjacent) & Der Platz \textbf{neben} mir ist frei. \\
    &       &       & The seat next to me is free. \\
    \hline
    % \pagebreak
    \SetCell[r=6]{m} über & \SetCell[r=6]{m} A & \SetCell[r=6]{m} over, above & Der Zug fuhr \textbf{über} die Brücke. \\*
    &       &       & The train went over the bridge. \\*
    \hline
    &       &       & Die Sonne hing \textbf{über} dem Meer \\*
    &       &       & The sun hung over the sea. \\*
    \hline
    &       &       & Wir redeten \textbf{über} den Film. \\*
    &       &       & We talked about the movie. \\*
    \hline
    \SetCell[r=8]{m} unter & \SetCell[r=2]{m} A & \SetCell[r=2]{m} under, below & Schau \textbf{unters} Bett vorm Schlafengehen \\
    &       &       & Look under the bed before going to sleep. \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} under, below & Ein Monster versteckt sich \textbf{unterm} Bett. \\
    &       &       & A monster is hiding under the bed. \\
    \hline
    & \SetCell[r=4]{m} D & \SetCell[r=4]{m} among & ein Schaf \textbf{unter} Wölfen \\
    &       &       & a sheep among wolves \\
    \hline
    &       &       & \textbf{unter} anderem (u.a.) \\
    &       &       & among other things / inter alia \\
    \hline
    \SetCell[r=8]{m} vor & \SetCell[r=2]{m} A & \SetCell[r=2]{m} in front of & Ich stellte mich \textbf{vors} Mikro(fon). \\
    &       &       & I went up to the microphone. \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} in front of & Der Tisch steht \textbf{vor} dem Fenster. \\
    &       &       & The table is in front of the window. \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} of & Angst \textbf{vor} Höhe \\
    &       &       & fear of heights \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} before & Ich muss es \textbf{vorm} Wochenende machen \\
    &       &       & I have to do it before the weekend. \\
    \hline
    \SetCell[r=5]{m} zwischen & \SetCell[r=2]{m} A & \SetCell[r=2]{m} between & etwas \textbf{zwischen} die Finger bekommen \\
    &       &       & to get a hold of something \textcolor{codegray}{[between one's fingers]} \\
    % &       &       & \\
    \hline
    & \SetCell[r=3]{m} D & \SetCell[r=3]{m} between & \textbf{zwischen} den Jahren \\
    &       &       & between the years \\
    &       &       & \textcolor{codegray}{[German expression for the time between Christmas and New Year's]}\\
    \hline
    \SetCell[r=2]{m} (an)statt & \SetCell[r=2]{m} G & \SetCell[r=2]{m} instead of & Statt der erwarteten Beförderung bekam er die Kündigung. \\
    &       &       & Instead of the expected promotion, he was let go. \\
    \hline
    \SetCell[r=2]{m} abzüglich & \SetCell[r=2]{m} G & \SetCell[r=2]{m} less, minus & \textbf{abzüglich} der Unkosten... \\
    &       &       & after expenses... \\
    \hline
    \SetCell[r=2]{m} angesichts & \SetCell[r=2]{m} G & \SetCell[r=2]{m} in view of, given & \textbf{Angesichts} der Tatsache, dass… \\
    &       &       & In view of the fact that… \\
    \hline
    % \pagebreak
    \SetCell[r=3]{m} außerhalb & \SetCell[r=3]{m} G & \SetCell[r=3]{m} outside of & \textbf{außerhalb} der Stadt \\*
    &       &       & outside the city \\*
    &       &       & \textcolor{codegray}{[famous cabaret in Berlin]}\\*
    \hline
    \SetCell[r=2]{m} bar & \SetCell[r=2]{m} G & \SetCell[r=2]{m} devoid of, barring any & \textbf{Bar} jeder Vernunft \\
    &       &       & against all reason \\
    \hline
    \SetCell[r=2]{m} bezüglich & \SetCell[r=2]{m} G & \SetCell[r=2]{m} concerning, regarding, with respect to & \textbf{bezüglich} unseres Telefonats \\
    &       &       & regarding our phone conversation \\
    \hline
    \SetCell[r=3]{m} binnen & \SetCell[r=3]{m} G & \SetCell[r=3]{m} within [a time period] & \textbf{binnen} einer Woche \\
    &       &       & within a/one week \\
    &       &       & \textcolor{codegray}{[also frequently used in the dative]}\\
    \hline
    \SetCell[r=2]{m} dank & \SetCell[r=2]{m} G & \SetCell[r=2]{m} thanks to & \textbf{dank} Ihrer Hilfe... \\
    &       &       & thanks to your help... \\
    \hline
    \SetCell[r=2]{m} einschließlich & \SetCell[r=2]{m} G & \SetCell[r=2]{m} including & \textbf{einschließlich} aller Kosten \\
    &       &       & all expenses included \\
    \hline
    \SetCell[r=2]{m} infolge & \SetCell[r=2]{m} G & \SetCell[r=2]{m} due to, because of & \textbf{infolge} einer schweren Erkrankung \\
    &       &       & due to a serious illness \\
    \hline
    \SetCell[r=2]{m} inmitten & \SetCell[r=2]{m} G & \SetCell[r=2]{m} in the middle/midst of & \textbf{inmitten} der Menge \\
    &       &       & in the middle of the crowd \\
    \hline
    \SetCell[r=3]{m} innerhalb & \SetCell[r=3]{m} G & \SetCell[r=3]{m} inside of & \textbf{innerhalb} eines kleinen Kreises \\
    &       &       & within a small circle \\
    &       &       & \textcolor{codegray}{[of people]}\\
    \hline
    \SetCell[r=3]{m} mangels & \SetCell[r=3]{m} G & \SetCell[r=3]{m} for lack of (esp. legal) & Das Verfahren war \textbf{mangels} Beweisen eingestellt. \\
    &       &       & The case was dismissed for lack of evidence. \\
    &       &       & \textcolor{codegray}{[assembly instructions]}\\
    \hline
    \SetCell[r=2]{m} mittels & \SetCell[r=2]{m} G & \SetCell[r=2]{m} by means of (esp. technical) & \textbf{mittels} eines Schraubenziehers... \\
    &       &       & using a screwdriver... \\
    \hline
    \SetCell[r=2]{m} oberhalb & \SetCell[r=2]{m} G & \SetCell[r=2]{m} above, beyond & \textbf{oberhalb} des Gefrierpunktes \\
    &       &       & above the freezing point \\
    \hline
    \SetCell[r=4]{m} unbeschadet & \SetCell[r=4]{m} G & \SetCell[r=4]{m} notwithstanding & \textbf{unbeschadet} des zuvor Gesagten... \\
    &       &       & despite/notwithstanding what was just said... \\
    \hline
    &       &       & \textbf{unbeschadet} dessen, dass... \\
    &       &       & notwithstanding the fact that... \\
    \hline
    \SetCell[r=2]{m} ungeachtet & \SetCell[r=2]{m} G & \SetCell[r=2]{m} in spite of & \textbf{Ungeachtet} der Drohungen des Irans will Deutschland Sanktionen. \\
    &       &       & In spite of Iran's threats, Germany wants sanctions. \\
    \hline
    \SetCell[r=2]{m} unterhalb & \SetCell[r=2]{m} G & \SetCell[r=2]{m} below, underneath & \textbf{unterhalb} der Armutsgrenze \\
    &       &       & below the poverty line \\
    \hline
    \SetCell[r=2]{m} unweit & \SetCell[r=2]{m} G & \SetCell[r=2]{m} not far from & \textbf{unweit} des Hotels \\
    &       &       & not far from the hotel \\
    \hline
    \SetCell[r=2]{m} zugunsten & \SetCell[r=2]{m} G & \SetCell[r=2]{m} in favor of, in the interests of & \textbf{zugunsten} meines Mandanten \\
    &       &       & in the interest of my client \\
    \hline
    \SetCell[r=5]{m} trotz & \SetCell[r=2]{m} G & \SetCell[r=2]{m} despite & \textbf{trotz} des schlechten Wetters \\
    &       &       & despite the bad weather \\
    \hline
    & \SetCell[r=2]{m} D & \SetCell[r=2]{m} despite & \textbf{trotz} dem schlechten Wetter \\
    &       &       & despite the bad weather \\
    \hline
    & \SetCell[c=3]{l,m,\linewidth-3cm} Native speakers disagree about whether the correct usage of "trotz" is genitive or dative. This is partly a North/South thing and partly a function of how it sounds with different nouns. We had to put it somewhere, but you can absolutely use it either way.\\
    \hline
    \SetCell[r=2]{m} während & \SetCell[r=2]{m} G & \SetCell[r=2]{m} during & \textbf{Während} des Semesters gelten folgende Sprechstunden... \\
    &       &       & During the semester the following office hours apply... \\
    \hline
    \SetCell[r=3]{m} wegen & \SetCell[r=3]{m} G & \SetCell[r=3]{m} because of & \textbf{wegen} Krankheit geschlossen \\
    &       &       & closed due to illness \\
    &       &       & \textcolor{codegray}{[Like "dank" and "trotz" this is very commonly used in the dative.]}
\end{longtblr}